---
title       : Exchange Rates and Exchange Rate Variability
subtitle    : Exchange Rate Visualizations
author      : Bernie Pilgram, PhD
job         : Nonlinear Dynamics & Predictive Analytics
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : [bootstrap, quiz, shiny, interactive]   # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

### Introduction 
---
## Determinants of exchange rate 
---
### __Economic factors__

- Economic policy, controlled and set by government agencies and central banks. 
- Economic conditions, e.g., balance of trade, inflation, economic growth and health, productivity of an economy

### __Political conditions__

- Internal, regional, and international political circumstances and events, e.g., destabilization of the political situation in Ukraine can negatively affect the value of its currency.

### __Market psychology__

- Market psychology and trader perceptions, e.g., long-term trends, technical trading considerations, "Buy the rumor, sell the fact". 

--- .class #id 

### Exchange rate visualization
---
## Can I predict exchange rates?
---
### __Let us check this [Shiny App](http://berniepilgram.shinyapps.io/CourseProjects_FXRate/) on exchange rates__
```{r FX_settings,echo=FALSE,message=FALSE,warning=FALSE}
# Preset variables for currency exchange rate calculations
# Load libraries required
library(quantmod); library(xts)

# Define variables
currencyA <- "USD"; currencyB <- "EUR"; currencyA_value <- 10000; days_past <- 50
```
- Choose from a set of currencies you want to exchange and check how the currency exchange rate changes over time, for example: `r format(currencyA_value,big.mark=",",scientific=F)` USD -> EUR at closing date `r format(Sys.Date()-1, format="%B %d, %Y")` and 
`r as.integer(days_past)` days prior.

```{r FX_over_time,fig.width=14,fig.height=5,fig.path='figures/',echo=FALSE,message=FALSE,warning=FALSE}
# Create function for currency conversion using the getFX() function from the
# quantmod package. Output is a single conversion for a certain currency pair at day
FXconversion <- function(currencyA,currencyB,amount,days){
        currency_ratio <- paste(currencyA,"/",currencyB,sep="")
        FXrate <- xts(get(getFX(currency_ratio,from = Sys.Date()-days,
                                to = Sys.Date()-days)))
        amount_converted <- as.numeric(amount*FXrate)
        return(amount_converted)}

# Create history of currency exchange rates or respectively exchanged currencies
FXhistory <- function(currencyA,currencyB,amount,days){
        currency_ratio <- paste(currencyA,"/",currencyB,sep="")
        FXrate <- xts(get(getFX(currency_ratio,from = Sys.Date()-days,
                                to = Sys.Date()-1)))
        amount_history <- amount*FXrate
        return(amount_history)}

# Calculate amount exchanged
currencyB_value_current <- FXconversion(currencyA,currencyB,currencyA_value,1)
# Calculate amount exchanged in the past
currencyB_value_past <- FXconversion(currencyA,currencyB,currencyA_value,days_past+1)
# Cacluate exchange rate over time
FX_over_time <- FXhistory(currencyA,currencyB,currencyA_value,days_past+2)

# Prepare for plot
FX_over_tmp <- FX_over_time; index <- 2:(length(FX_over_tmp)-1); index[1]<-1
FX_over_time_line <- FX_over_tmp[-index]
plot(FX_over_time,main = "Amount as Value of Currency over Time", ylab = currencyA)
points(FX_over_time[2],pch=19); points(FX_over_time[length(FX_over_time)],pch=19)
lines(FX_over_time_line,pch=23,lty=2,col="red"); legend("topleft","Current and past date",
       pch=19, title = "Values at closing dates")
```

### Ready to join the team?

--- &radio
### Exchange rates over time - Are you ready to join the team?
---
### USD/EUR exchange rate on `r format(Sys.Date()-days_past-1, format="%B %d, %Y")`
---
How much money did you loose when you changed `r format(currencyA_value,big.mark=",",scientific=F)` USD to EUR on `r format(Sys.Date()-days_past-1, format="%B %d, %Y")` compared to closing date `r format(Sys.Date()-1, format="%B %d, %Y")`?

1. 100.01 USD
2. _`r as.double(currencyB_value_current)-as.double(currencyB_value_past)`_ USD
3. 76.34 USD
*** .hint
Please check the previous slide!
*** .explanation
Are you ready to join the team?

--- .class #id 

### Research Project and Capstone Project
---
## Modeling of exchange rate behaviour 
---
### 1. Review of current theories and models

- International parity conditions, e.g., Relative Purchasing Power Parity
- Balance of payments model, Asset market model, algorithmic trading models

### 2. Development of an exchange-rate model

- Regression analysis (linear, non-linear) of economic factors and variables to model economic factors, a combination of both classical and Bayesian models will be considered.  
- Strategic Reasoning to model (1) political conditions and (2) market psychology

### 3. Deployment of model at webserver

- User friendly setup: User can choose from a set of economic factors and variables, and from a set of decision points for Strategic Reasoning







